Appendix

### 2025-03-27 manuscript_rmarkdown/manuscript_appendix.tex
WIP on simulations with clean data.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{amsmath}
\usepackage{booktabs}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\sgn}{sgn}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Appendix},
  pdfauthor={Alex Zwanenburg, Steffen Löck},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-03-13}

@@

@@

Applying conventional and invariant Box-Cox transformations to age of
patients with lung cancer (Loprinzi et al. 1994) yielded the following
results: no transformation (sum of residuals with normal distribution
\(\sum r_i = 16.5\)); conventional transformation (\(\lambda = 1.9\),
\(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
\(\sigma_{BC} = 0.4 \cdot 10^3\)); Raymaekers and Rousseeuw's robust
adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\));
location- and scale-invariant transformation (\(\lambda = 1.7\),
\(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\), \(\sigma_{BC} = 0.8\)); and
robust location- and scale-invariant transformation (\(\lambda = 1.5\),
\(\sum r_i = 11.6\), \(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\)).

@@

Applying conventional and invariant Box-Cox transformations to the body
mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
following results: no transformation (residual sum \(\sum r_i = 48.0\));
conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); location- and
scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 27.3\),
\(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\)); and robust location- and
scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 25.2\),
\(\mu_{BC} = 0.4\), \(\sigma_{BC} = 0.6\)).

@@

Applying conventional and invariant Box-Cox transformations to the
latitude of houses in the Ames housing dataset (De Cock 2011) yielded
the following results: no transformation (residual sum
\(\sum r_i = 328\)); conventional transformation (\(\lambda = 62.1\),
\(\sum r_i = 319\), \(\mu_{BC} = 1.1 \cdot 10^{99}\),
\(\sigma_{BC} = 0.0 \cdot 10^{99}\)); Raymaekers and Rousseeuw's robust
adaptation (\(\lambda = 96.0\), \(\sum r_i = 319\),
\(\mu_{BC} = 6.2 \cdot 10^{153}\),
\(\sigma_{BC} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
transformation (\(\lambda = 1.9\), \(\sum r_i = 312\),
\(\mu_{BC} = 2.3\), \(\sigma_{BC} = 0.9\)); and robust location- and
scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\)).

@@

The Top Gear dataset contains data on 297 vehicles, with outliers
related highly fuel-efficient vehicles (Alfons 2021). Applying
conventional and invariant Box-Cox transformations to the fuel
consumption feature yielded the following results: no transformation
(residual sum \(\sum r_i = 54\), \(p=0.76\)); conventional
transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.55\)); location- and
scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p=0.02\)); and robust
location- and scale-invariant transformation (\(\lambda = 1.1\),
\(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
\(p=0.83\)).

@@

The ischemic stroke dataset contains historic data from 126 patients
with risk at ischemic stroke (Kuhn and Johnson 2019). Applying
conventional and invariant Box-Cox transformations to the maximum
arterial wall thickness feature yielded the following results: no
transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
\(\mu_{BC} = 1.0\), \(\sigma_{BC} = 0.2\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.60\)); location- and
scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.01\)); and robust
location- and scale-invariant transformation (\(\lambda = 0.5\),
\(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
\(p=0.35\)).

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n}$ for
empirical central normality at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The presented values should be divided by $100$.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \ $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\

\midrule
    5 & 24.02 & 21.26 & 17.37 & 14.02 & 12.04 &  9.57 &  7.64 &  5.54 & 2.35 \\
   10 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 &  7.65 &  6.79 &  5.87 & 4.23 \\
   20 & 15.12 & 13.17 & 10.03 &  7.74 &  6.80 &  6.13 &  5.58 &  5.04 & 4.07 \\
   50 &  9.99 &  8.70 &  6.69 &  5.26 &  4.67 &  4.26 &  3.92 &  3.59 & 3.02 \\
  100 &  7.13 &  6.21 &  4.83 &  3.79 &  3.38 &  3.08 &  2.87 &  2.63 & 2.24 \\ 
  200 &  5.19 &  4.54 &  3.51 &  2.77 &  2.47 &  2.26 &  2.09 &  1.92 & 1.65 \\ 
  500 &  3.47 &  3.06 &  2.38 &  1.86 &  1.65 &  1.51 &  1.40 &  1.28 & 1.08 \\
 1000 &  2.65 &  2.35 &  1.84 &  1.44 &  1.27 &  1.16 &  1.07 &  0.98 & 0.81 \\
 2000 &  2.13 &  1.91 &  1.53 &  1.20 &  1.06 &  0.96 &  0.88 &  0.79 & 0.65 \\
 5000 &  1.69 &  1.56 &  1.30 &  1.07 &  0.95 &  0.87 &  0.80 &  0.72 & 0.58 \\
10000 &  1.50 &  1.41 &  1.23 &  1.05 &  0.96 &  0.89 &  0.83 &  0.77 & 0.64 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:normalisation-before-transformation-appendix}
\begin{tabular}{l | c c c c c}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{multirow}
\newtheorem{definition}{Definition}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\sgn}{sgn}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Appendix},
  pdfauthor={Alex Zwanenburg, Steffen Löck},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-03-27}

@@

\section{Appendix D: Simulations on clean
data}\label{appendix-d-simulations-on-clean-data}

{[}Generated data: clean, n = 30, n = 100, n = 1000, lambda = 0.1, 1.0,
1.9{]}

\begin{table}
\begin{center}
\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-residuals-appendix}
\small{
\begin{tabular}{l | l r r r r r r r r r}

\toprule
& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\

\midrule

none                                  & & 4.0 & 3.4 & 4.6 & 40.1 & 8.9 & 20.8 & 184.6 & 10.1 & 191.8 \\
conventional                          & & 2.7 & 3.3 & 3.6 &  6.5 & 8.1 &  5.6 &  14.2 & 10.1 &  16.1 \\
conventional (z-score norm.)          & & 2.5 & 3.3 & 3.7 &  9.8 & 8.1 &  5.6 &  24.1 & 10.1 &  15.9 \\
conventional (robust scaling)         & & 2.6 & 3.3 & 3.7 &  7.4 & 8.2 &  5.6 &  15.5 & 10.1 &  14.0 \\
Raymaekers-Rousseeuw                  & & 3.8 & 3.3 & 3.6 &  6.5 & 8.5 &  5.6 &  14.4 & 11.0 &  17.6 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 2.6 & 3.3 & 3.7 &  9.8 & 8.5 &  5.6 &  24.1 & 11.1 &  15.9 \\
Raymaekers-Rousseeuw (robust scaling) & & 2.6 & 3.3 & 3.7 &  7.4 & 8.6 &  5.6 &  15.4 & 11.1 &  15.1 \\
invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 & 8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
robust invariant                      & & 3.0 & 3.3 & 3.6 &  5.2 & 8.3 &  5.9 &  19.9 & 13.6 &  32.9 \\
\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-lambda-appendix}
\small{
\begin{tabular}{l | l r r r r r r r r r}

\toprule
& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\

\midrule

conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
robust invariant                      & & 0.4 & 1.3 & 1.5 &  0.0 & 0.8 & 1.7 &  0.0 & 1.1 & 2.0 \\
\bottomrule
\end{tabular}
}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{
p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:clean-transformation-p-value-appendix}
\small{
\begin{tabular}{p{3.5cm} | l r r r r r r r r r}

\toprule
& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\

\midrule

none                                  & & 0.986 & 0.614 & 0.323 & $<0.001$ & 0.063 & 0.016 & $<0.001$ & 0.997 & $<0.001$ \\
conventional                          & & 0.986 & 0.464 & 0.488 &   0.675  & 0.139 & 0.620 &   0.902  & 0.999 &   0.698 \\
conventional (z-score norm.)          & & 0.979 & 0.440 & 0.477 &   0.200  & 0.141 & 0.770 &   0.419  & 0.999 &   0.723 \\
conventional (robust scaling)         & & 0.982 & 0.439 & 0.484 &   0.487  & 0.130 & 0.683 &   0.883  & 0.999 &   0.817 \\
Raymaekers-Rousseeuw                  & & 0.989 & 0.464 & 0.488 &   0.675  & 0.446 & 0.620 &   0.876  & 0.992 &   0.821 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 0.992 & 0.440 & 0.477 &   0.200  & 0.448 & 0.770 &   0.460  & 0.991 &   0.719 \\
Raymaekers-Rousseeuw (robust scaling) & & 0.993 & 0.439 & 0.484 &   0.487  & 0.422 & 0.683 &   0.887  & 0.991 &   0.848 \\
invariant                             & & 0.993 & 0.355 & 0.461 &   0.720  & 0.153 & 0.490 &   0.910  & 0.999 &   0.682 \\
robust invariant                      & & 0.994 & 0.407 & 0.484 &   0.642  & 0.291 & 0.489 &   0.848  & 0.996 &   0.628 \\
\bottomrule
\end{tabular}
}
\end{center}
\end{table}

@@

Applying conventional and invariant Box-Cox transformations to age of
patients with lung cancer (Loprinzi et al. 1994) yielded the following
results: no transformation (sum of residuals with normal distribution
\(\sum r_i = 16.5\), \(p = 0.71\)); conventional transformation
(\(\lambda = 1.9\), \(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
\(\sigma_{BC} = 0.4 \cdot 10^3\), \(p = 0.94\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\),
\(p = 0.94\)); location- and scale-invariant transformation
(\(\lambda = 1.7\), \(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\),
\(\sigma_{BC} = 0.8\), \(p = 0.94\)); and robust location- and
scale-invariant transformation (\(\lambda = 1.5\), \(\sum r_i = 11.6\),
\(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\), \(p = 0.93\)).

@@

Applying conventional and invariant Box-Cox transformations to the body
mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
following results: no transformation (residual sum \(\sum r_i = 48.0\),
\(p < 0.001\)); conventional transformation (\(\lambda = -0.5\),
\(\sum r_i = 32.2\), \(\mu_{BC} = 2.1\),
\(\sigma_{BC} = 4 \cdot 10^{-3}\), \(p = 0.06\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\), \(p = 0.06\));
location- and scale-invariant transformation (\(\lambda = 0.5\),
\(\sum r_i = 27.3\), \(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\),
\(p = 0.10\)); and robust location- and scale-invariant transformation
(\(\lambda = 0.2\), \(\sum r_i = 25.2\), \(\mu_{BC} = 0.4\),
\(\sigma_{BC} = 0.6\), \(p = 0.44\)).

@@

Applying conventional and invariant Box-Cox transformations to the
latitude of houses in the Ames housing dataset (De Cock 2011) yielded
the following results: no transformation (residual sum
\(\sum r_i = 328\), \(p < 0.001\)); conventional transformation
(\(\lambda = 62.1\), \(\sum r_i = 319\),
\(\mu_{BC} = 1.1 \cdot 10^{99}\), \(\sigma_{BC} = 0.0 \cdot 10^{99}\),
\(p < 0.001\)); Raymaekers and Rousseeuw's robust adaptation
(\(\lambda = 96.0\), \(\sum r_i = 319\),
\(\mu_{BC} = 6.2 \cdot 10^{153}\), \(\sigma_{BC} = 0.3 \cdot 10^{153}\),
\(p < 0.001\)); location- and scale-invariant transformation
(\(\lambda = 1.9\), \(\sum r_i = 312\), \(\mu_{BC} = 2.3\),
\(\sigma_{BC} = 0.9\), \(p < 0.001\)); and robust location- and
scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\), \(p < 0.001\)).

@@

The Top Gear dataset contains data on 297 vehicles, with outliers
related highly fuel-efficient vehicles (Alfons 2021). Applying
conventional and invariant Box-Cox transformations to the fuel
consumption feature yielded the following results: no transformation
(residual sum \(\sum r_i = 54\), \(p=0.58\)); conventional
transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p < 0.001\)); Raymaekers
and Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.32\)); location- and
scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p < 0.001\)); and robust
location- and scale-invariant transformation (\(\lambda = 1.1\),
\(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
\(p=0.67\)).

@@

The ischemic stroke dataset contains historic data from 126 patients
with risk at ischemic stroke (Kuhn and Johnson 2019). Applying
conventional and invariant Box-Cox transformations to the maximum
arterial wall thickness feature yielded the following results: no
transformation (residual sum \(\sum r_i = 110\), \(p=0.86\));
conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
\(\mu_{BC} = 1.0\), \(\sigma_{BC} = 0.2\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.90\)); location- and
scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.02\)); and robust
location- and scale-invariant transformation (\(\lambda = 0.5\),
\(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
\(p=0.60\)).

@@

Critical values for the central normality test and its empirical variant
are found in Table \ref{tab:central-normality-critical-statistic} and
Table \ref{tab:empirical-central-normality-critical-statistic},
respectively.

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\

\midrule
5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\

\midrule
5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-critical-statistics-1}

\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.70}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$. The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e. with fully normal data where 10\% of elements are replaced by outliers. emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
\end{figure}

@@

\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:normalisation-before-transformation-appendix}
\begin{tabular}{l | r r r r r}
